Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-06-09 LaTex/paper.tex
more literature and explanations everywhere

%\title{Prediction and Evaluation of individual Probability Density Functions with Machine Learning}
\title{Demand Forecasting of individual Probability Density Functions with Machine Learning}

@@

\section{Introduction}

Demand forecasting is one of the main challenges for retailers and at the core of business
operations. Due to its stochastic nature, demand is difficult to forecast as it depends on
many influencing factors and the realized demand can be interpreted as a random variable
that is described by an appropriate probability density function (PDF). In order to make
operational decisions, an optimal point estimator has to be defined that  can be used to
derive ordering decisions used in the replenishment process of the retailer. Demand
estimation is further complicated by the fact that retailers typically only observed
realized sales and not the actual demand, in case the demand exceeds the current stock
level, the data become censored. This decision process is complicated by a range of
factors: Even in the case of accurate  demand forecasts, the decision maker has to balance
conflicting metrics to reach an optimal decision: Ordering to few items may result in
stockout situations resulting in unrealized demand and unsatisfied customers. Ordering too
many items results in excess inventory which increases transport and storage costs and, in
the case of perishable goods, excessive waste as spoilt items need to be disposed of at
additional cost in addition to  the environmental impact. This situation is particularly
noticeable in the so called `'ultra-fresh'' category which includes items such as bakery
products, ready-meals, fresh diary products or certain meat products such as ground meat.
These items typically have a shelf-life from less than a business day to a few business
days at most with a continuous spectrum  depending on the exact item. In many situations,
additional constraints have to be considered to reach an optimal ordering decision:
Delivery cycles of items may vary depending on the type of item and the wholesaler or
manufacturer from which  they are procured. Retailers also operate at a given service
level to guarantee that a certain level of demand can be fulfilled. The exact service
level typically depends on the overall business policy of the retailer and may also depend
on individual products, ranging from ``never-out-of-stock'' items to a service level
exceeding e.g. 90\%.

The availability of Big Data allows capturing, storing and processing a vast amount of
data associated with demand such as historic sales records,  information about promotional
events or advertisements, pricing information, local weather at retail locations, seasonal
information  as well as a wide range of further variables. Modern machine learning
algorithms can then be used to predict the per-item demand distribution, corrected for
censored data from which an optimal point estimator can be derived to be used in the
subsequent ordering decision. It is important to note that demand as a random variable is
not identically and independently distributed (i.i.d.): While the probability distribution
describing the demand can be attributed to a given family or parametrization, the exact
parameters vary: Seasonal effects, finite life cycles of products and the introduction of
new products influence the demand distribution, as well as the local weather at the retail
location or the retail location itself in terms of size, assortment range, customer
diversity and other factors. The retailers themselves also actively influence demand by
using advertisements to highlight products, offering rebates or discounts for specific
products as well as pursuing an active pricing strategy. This means that while we can
generally assume that demand follows a specific type of probabilty distribution, its
parameters are unique to the  instance for which an estimate is required. For example, the
probability distribution governing the demand of a particular item is specific to the
item, date and retail location for which the forecast is made and depends on a wide range
of further influencing factors.

The remainder of the paper is organized as follows:
We first review the relevant literature and existing work in sec. \ref{sec:LitRev}.
We then describe our method to predict individual negative binomial PDFs by means of a
parametric approach including two distinct machine learning models for mean of variance in
sec. \ref{sec:pdfEstimation}.
After that we describe methods for the qualitative and quantitative evaluation of PDF
predictions in sec. \ref{sec:pdfEvaluation}.
Finally, we present a demand forecasting example to show an application of our methods in
sec. \ref{sec:example}.

@@

Inventory management offers a rich theory and the extensive body of research can be
broadly grouped into the following two categories where the inventory control problem is
either based on some knowledge of the underlying demand  distribution or an integrated
approach that seeks to map directly from the available data (historic sales records and 
further variables) to the ordering decision. The latter approach is taken e.g. in
\cite{beutel2012safety,ban2019big,bertsimas2020predictive} and aims to avoid estimating
the underlying probability distribution. Although this approach seems preferable since it
avoids determining the full demand distribution and results directly in the desired
operational decision (order quantity), it faces several drawbacks. First of all, the full
probability distribution for the demand of an specific item at a given sales location and
business day includes all available information including the uncertainty of the modelled
demand. This can be used to simulate the performance on a per-item level and e.g. optimize
the impact on business strategy decisions on conflicting metrics such as stock out- and
waste-rate. Additionally, having the full demand probability distribution available allows
for an accute assessment of the forecast quality at all quantiles of the distribution,
including the often extensive tails of the distribution, as well as the analysis of
long-term  effect as the demand predictions including all future planned advertisements or
price-changes can be included into long-term forecasts. Furthermore, deriving the ordering
decision  directly couples the demand process with the complex delivery cycles and
constraints, keeping the steps separate allows greater operational flexibility and reduces
the impact of changing manufacturers or wholesalers, as well as  allowing a quick response
to changes in the delivery schedule wihtout having to re-compute the implicit demand
underlying the ordering decision. Finally, long-term demand forecasts may be shared with
other business units or external vendors  and wholesalers to ease their planning for the
production and supply-chain processes upstream of the retailer.

In contrast, more traditional inventory control systems rely on the knowledge of the
demand distribution in one form or another.  see e.g. \cite{silver1998} for an overview.
In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored
at regular intervals and orders are dispatched once the inventory level reaches a minimal
value $s$. In case of linear holding and shortage costs, such policies are optimal
\cite{Scarf1959}, although perishable goods pose more challenges, see e.g.
\cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be
included in these kind of inventory control systems \cite{minner2010periodic}. Perishable
goods are well described by the `'newsvendor-problem'' \cite{Edgeworth} where in the
simplest case all stock perishes at the end of the selling period (e.g. a business day). 
For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming
linear underage and overage costs $b,h >0$, the optimal quantile
$q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated
exactly. The main objective in any of these approaches is to determine the underlying
demand distribution. The simplest approach is to just use the observed sales events and
forecast these as a time-series  (see e.g. \cite{alwan2016}) or via sample average
approximation (SAA). (see e.g. \cite{shapiro2014} for an overview over SAA). However,
these approaches do not make use of any data apart from the sales record themselves,
although we know that many variables such as price, advertisements, etc. are highly
correlated with demand. Additionally, it is critical that we indeed reconstruct a demand
distribution, hence a simple point-estimator as provided by the most common statistical
techniques and machine-learning approaches will not suffice. Additionally, demand is not
identically and independently distributed (i.i.d.) but depends not only on external
factors such as season, weather, product life-cycle, but is also actively changed by the
retailer by setting a specific price, offering rebates or running advertisements.
Additionally, the demand implicitly depend on the location of the retail outlet as well as
the specifics of that location such as product  assortment influencing the choice of
possible replacement articles and many more. These complications are the main reason we
cannot treat the replenishment process as $n$ independent newsvendor-type problems.
Instead, we need to determine the full demand distribution from data, conditional on the
relevant variables such as date, location and item, taking all auxiliary data such as
article charactersitics, pricing, advertisements, retail location details, etc. into
account. This can be done in several ways: First, we can use a neural network
\cite{Feindt2006190} to learn the distribution from data and return a full distribution
per item, store and day from which the relevant quantile can be estimated. Similarly,
using quantile regression \cite{koenker2001} this approach can be implemented in different
frameworks. Alternatively, one can assume a given demand model and fit the model
parameters instead of reconstructing the complete distribution. This approach is
computationally favourable, as fewer parameters need to be estimated compared to the case
of the full distribution. Empirically, one can determine the best fitting distribution
from data \cite{adan1995}. However, given the stochastic nature of the demand, such an
empirically determined distribution is not expected to be stable and prone to sudden
changes. Instead, the choice of the demand distribution should be motivated by theoretic
considerations. The discrete demand is typically modelled as a negative binomial
distribution (NBD), also known as Gamma-Poisson distribution
\cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This
distribution arises if the Poisson parameter $\mu$ is a random variable itself that
follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$ and
is over-dispersed compared to the Poisson distribution for which $\mu = \sigma^2$.

Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be
determined for each item at each sales location and ordering time, depending on all
auxiliary data describing article details, retail location and influencing factors such as
pricing and advertisement information.

\subsection*{Summary of Contributions}
What we are going to show.

@@

To predict an individual PDF by means of a parametric approach, one has to rely on a model
assumption about the underlying distribution of the random variable to be predicted. In
the following, we describe such a parametric approach with the model assumption of a
negative binomial probability distribution. For a PDF prediction under a negative binomial
model assumption, there is the need for the estimation of two parameters, for example its
mean and its variance.
In any case, as there is a strong correlation between mean and variance, it is beneficial to include the corresponding
predicted mean from the mean model as feature in the variance model.

This can be done using two independent models one,  to estimate the mean and the other
for the variance. At least in principle, any method can be used. However, as argued above,
in case of demand forecasting, each prediction is highly specific to the circumstances
in which it is used (such as product ID, day and store location) and may
depend on  a multitude of describing variables (features).
Machine learning algorithms are ideally suited for this task and in the
following we will use the ``Cyclic Boosting'' algorithm \cite{Wick2019}.
The major benefits of this algorithm are that it is not only extremly performant in practical 
applications of demand forecasting at large scale, but also fully explainable. Traditional machine
learning algorithms are typically ``black box'' approaches where the individual decision cannot be 
explained. Cyclic Boosting on the other hand allows to follow how each individual prediction was made.
The main idea of this algorithm is the following: When used for regression,
the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following
way:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The parameters $f^k_j$ are the model parameters that are determined from 
features $j$. Each feature is discretisized appropriately into $k$ bins to reflect the 
specific behaviour of the feature. The global mean $\mu$ is determined from all observed
target values $y$ observed in the data. The factors  $f^k_j$ are determined
iteratively until convergence is reached and regularization techniques are applied
to avoid overtraining and improve the generalization ability of the algorithm.
The relative strength of the factors  $f^k_j$ are then directly interpretable in
relation how important a specific feature is for each individual prediction where deviations
from  $f^k_j=1$ indicate high importance.

%One can achieve this with two subsequent, independent machine learning models, the first
%to estimate the mean and the second to estimate the variance. The features may or may not
%differ between the mean and variance estimation models.

The assigned mean and variance predictions can then be used to generate individual
probability density functions according to a functional negative binomial distribution
assumption for each sample.

%While any machine learning algorithm outputting a conditional mean as prediction may be
%used to get an approximate estimate for the mean of an underlying negative binomial
%distribution, it is preferable to choose an algorithm using a negative binomial assumption
%for its inner regularization methods. An example for such a machine learning algorithm is
%Cyclic Boosting in multiplicative regression mode .
%In the upcoming section, we describe a method to predict the variance of a negative
%binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.

@@

\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}
with mean $\mu$ and dispersion parameter $r$. The target variable $y$ takes the values $y = 0, 1, 2, ...$.

@@

As stated in \eqn \eqref{eqn:r} the values $r_i$ are defined by the Cyclic Boosting
model parameters $f^k_j$ for each feature $j$ and bin $k$. For any concrete observation
$i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the
subsequent look-up into which bin this observation falls. The dispersion parameter $r$ is
bound to the interval $[1, \infty]$.

\textcolor{red}{Hm, this part is a bit quick. We should add a sentence to the NB loss 
and how this is motivated and then how this is connected to \eqn \eqref{eqn:r} and how
\eqn \eqref {eqn:cb} from CB comes into play.}
\begin{equation} \label{eqn:r}
r_i = 1 + \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}

\textcolor{red}{Hm, this part is not quite clear - need to add some details what is meant}
As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an
iterative method corresponding to a cyclic coordinate descent, processing one feature with
all its bins at a time until convergence. Unlike in the original multiplicative regression
mode of Cyclic Boosting, the minimization of the loss function in \eqn
\eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method,
for example a random search. All other advantages of Cyclic Boosting presented in
\cite{cyclicboosting}, like for example individual explainability of predictions, remain
valid for its negative binomial width mode.

@@

Statistical or machine learning methods that predict full individual probability functions typically lack
quantitative, or at least qualitative, evaluation of the PDF model output to assess its
correctness. In the case of an estimation of the determining parameters of an assumed
functional form for the PDF, assessing the correctness of the PDF model output refers to
the evaluation of the accuracy of the prediction of the different determining parameters,
e.g. mean and variance of a negative binomial distribution, and the validation of the
model assumptions, i.e. the choice of the underlying PDF, by comparing the probability
density estimations to observed data. However, the two methods presented in this section,
namely cumulative distribution function (CDF) histograms and inverse quantile plots, are
neither restricted to the evaluation of parametric PDF estimations nor to a specific
functional form, like a negative binomial, but can be used generically.

@@

\subsubsection{CDF Histogram}
\label{sec:cdf_histo}

\textcolor{red}{need to explain intuition a bit more, a diagram would be helpful.}
The idea here is that the individual PDF predictions are first transformed into individual
CDF predictions and the CDF values, also known as quantiles, of the corresponding truth
values are then filled to a histogram.

\textcolor{red}{need to explain a bit more what we see and why we expect a uniform distribution.}
Figure \eqref{fig:cdf_histos} illustrates five different CDF histograms. If the PDF
prediction is correct, a flat uniform distribution should occur for the CDF histogram,
like illustrated by the final far-right column of shaded circles.

In the example of a parametric negative binomial PDF estimation of section
\eqref{sec:pdfEstimation}, this would indicate that both the mean and the variance
estimations as well as the choice of a negative binomial distribution as underlying PDF
were correct. Conversely, the first two plots of \fig \ref{fig:cdf_histos} indicate a
bias in the variance estimation and the third and fourth plot indicate a bias in the mean
estimation.
\textcolor{red}{Need to explain why we see that. Would be good to show the respective distributions
(or at least one example) of wher it doesn't fit.}

@@

@@

... predict an individual demand volume for different product-location-date combinations ...
... example showing full prediction and evaluation chain ...

... plot showing individual PDF example ...

@@

... plot showing cdf histo ...

@@

... plot showing invquant profile for mean prediction on X-axis ...

@@

... plot showing invquant profile for product groups on X-axis ...

@@

...

>>>>

\title{Demand Forecasting of individual Probability Density Functions with Machine Learning}

@@

PDF
\section{Introduction}

Demand forecasting is one of the main challenges for retailers and at the core of business operations. Due to its stochastic nature, demand is difficult to forecast as it depends on many influencing factors and the realized demand can be interpreted as a random variable that is described by an appropriate probability density function (PDF). In order to make operational decisions, an optimal point estimator has to be defined, that can be used to derive ordering decisions used in the replenishment process of the retailer. Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored. The ordering decision process is complicated by a range of factors: Even in the case of accurate  demand forecasts, the decision maker has to balance conflicting metrics to reach an optimal decision: Ordering to few items may result in stockout situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life of less than a business day to a few business days at most, with a continuous spectrum depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from "never-out-of-stock'' items to a service level exceeding e.g. 90\%.

The availability of Big Data allows capturing, storing and processing a vast amount of data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information as well as a wide range of further variables. Modern machine learning algorithms can then be used to predict the per-item demand distribution, corrected for censored data, from which an optimal point estimator can be derived to be used in the subsequent ordering decision. It is important to note that demand as a random variable is not identically and independently distributed (i.i.d.). While the probability distribution describing the demand can be attributed to a given family or parametrization, the exact parameters vary: Seasonal effects, finite life cycles of products and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity and other factors. The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy. This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the  instance for which an estimate is required. For example, the probability distribution governing the demand of a particular item is specific to the item, date and retail location for which the forecast is made and depends on a wide range of further influencing factors.

The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec. \ref{sec:LitRev}. We then describe our method to predict individual negative binomial PDFs by means of a parametric approach including two distinct machine learning models for mean of variance in sec. \ref{sec:pdfEstimation}. After that we describe methods for the qualitative and quantitative evaluation of PDF predictions in sec. \ref{sec:pdfEvaluation}. And finally, we present a demand forecasting example to show an application of our methods in sec. \ref{sec:example}.

@@

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand  distribution or an integrated approach that seeks to map directly from the available data (historic sales records and further variables) to the ordering decision. This approach is often referred to as "data-driven newsvendor'' and discussed  e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution and use the available data to derive the operational decisions directly.  An overview of a range approaches can also be found in \cite{huber2019data}.

Although this approach seems preferable, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks. First of all, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand. This can be used to simulate the performance on a per-item level and e.g. optimize the impact on business strategy decisions on conflicting metrics such as stock out- and waste-rate. Additionally, having the full demand probability distribution available allows for an accurate assessment of the forecast quality of all quantiles of the distribution, including the often extensive tails of the distribution. Separating the demand forecast from the operational decisions (i.e. calculating the  order quantities for the next delivery cycle) also allows to evaluate longer-term planning or counterfactual analysis to investigate how the demand would have been had the input variables had different values. This way, effects of promotions, advertisements or price-changes can be analyzed  with the same model that is used for operational decisions. Furthermore, deriving the ordering decision directly couples the demand process with the complex delivery cycles and constraints, while keeping the steps separate allows greater operational flexibility and reduces the impact of changing manufacturers or wholesalers, as well as  allowing a quick response to changes in the delivery schedule without having to re-compute the implicit demand underlying the ordering decision. Finally, long-term demand forecasts may be shared with other business units or external vendors  and wholesalers to ease their planning for the production and supply-chain processes upstream of the retailer. From the perspective of industrial practice, modelling the demand separately from deriving the subsequent orders  has the additional benefit that multiple retail chains can benefit from any improvement in the model description even if the concrete retailers are unrelated to each other. For example, if a particular effect was identified at some retailer A and included in the machine learning model, the improvement can be rolled out immediately or on request to all other retailers using this system on a planetary scale without having to adapt the underlying machine learning model for each retailer individually.

In contrast to the direct mapping the observed data to ordering decisions, more traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g. \cite{silver1998} for an overview. In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$. In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g. \cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}. Perishable goods are well described by the "newsvendor-problem'' \cite{Edgeworth}, where in the simplest case all stock perishes at the end of the selling period (e.g. a business day). For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly.

The main objective in any of these approaches is to determine the underlying demand distribution. The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g. \cite{alwan2016}) or via sample average approximation (SAA), see e.g. \cite{shapiro2014} for an overview. However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price or advertisements influence, and therefore are highly correlated with, the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e. the  calculation of the optimal order quantity.

However, in order to be able to fully optimize the operational decision, it is critical that one indeed reconstructs a full demand distribution. This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine-learning approaches, will not suffice. Additionally, we need to consider that demand is not i.i.d., but depends on external factors such as season, weather, product life-cycle, and is also actively changed by the retailer by setting a specific price, offering rebates or running advertisements. Additionally, the demand implicitly depends on the location of the retail outlet as well as the specifics of that location, such as product  assortment influencing the choice of possible replacement articles and many more. These complications are the main reason we cannot treat the replenishment process as $n$ independent newsvendor-type problems.

Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc. into account. This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the  full probability distribution or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favourable, as fewer parameters need to be estimated compared to the case of the full distribution. Empirically, one can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modelled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself that follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution for which $\mu = \sigma^2$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

\subsection*{Summary of Contributions}

In the following, we will demonstrate how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer. Typically this means that the full demand distribution has to be estimated per SKU for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a "black-box'' machine learning model, using Cyclic Boosting allows to interpret how each individual prediction was made and what the most important variables are per individual prediction.

Additionally, we show how the cumulative distribution function (CDF) or inverse quantiles can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution. This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as  order quantities as well as strategic business decisions by the retailer.

@@

To predict an individual PDF by means of a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. In the following, we describe such a parametric approach with the model assumption of a negative binomial probability distribution. For a PDF prediction under a negative binomial model assumption, there is the need for the estimation of two parameters, for example its mean and its variance.

This can be done using two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, as argued above, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as product ID, date, and store location) and may depend on  a multitude of describing variables (features). Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm. The major benefits of this method are that it is not only extremely performant in practical applications of demand forecasting at large scale, but also fully explainable. Traditional machine learning algorithms are typically "black box'' approaches, where the individual decision cannot be explained. Cyclic Boosting on the other hand allows to follow how each individual prediction was made. The main idea of this algorithm is the following: When used for regression, the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following way:

\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}

The parameters $f^k_j$ are the model parameters that are determined from features $j$. Each feature is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature. The global mean $\mu$ is determined from all target values $y$ observed in the data. The factors $f^k_j$ are determined iteratively until convergence is reached and regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The relative strength of the factors  $f^k_j$ are then directly interpretable in relation to how important a specific feature is for each individual prediction, where deviations from  $f^k_j=1$ indicate high importance.

\textcolor{red}{Need for some details on Cyclic Boosting training process}

For the PDF estimation, two subsequent, independent Cyclic Boosting models are used, the first to estimate the mean and the second to estimate the variance. The features may or may not differ between the mean and variance estimation models. But in any case, as there is a strong correlation between mean and variance, it is beneficial to include the corresponding predicted mean from the mean model as feature in the variance model. The assigned mean and variance predictions can then be used to generate individual PDFs according to a functional negative binomial distribution assumption for each sample.

While any machine learning algorithm outputting a conditional mean as prediction may be used to get an approximate estimate for the mean of an underlying negative binomial distribution, it is preferable to choose an algorithm using a negative binomial assumption for its inner regularization methods, like Cyclic Boosting in multiplicative regression mode.

In the upcoming section, we describe a method to predict the variance of a negative binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.

@@

\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}

with mean $\mu$ and dispersion parameter $r$. The target variable $y$ takes the values $y = 0, 1, 2, ...$.

@@

As stated in \eqn \eqref{eqn:r}, the values $r_i$ are defined by the Cyclic Boosting model parameters $f^k_j$ for each feature $j$ and bin $k$. For any concrete observation $i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the subsequent look-up into which bin this observation falls. The dispersion parameter $r$ is bound to the interval $[1, \infty]$.

\begin{equation} \label{eqn:r}
r_i = 1 + \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}

\textcolor{red}{Need for more details.}

As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an iterative method corresponding to a cyclic coordinate descent, processing one feature with all its bins at a time until convergence. Unlike in the original multiplicative regression mode of Cyclic Boosting, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method, for example a random search. All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.

@@

Many statistical and most machine learning methods do not provide a full probability density distribution as result. Instead, these methods typically predict a single numerical quantity that is then compared to the observed concrete realization of the random variable using metrics such as the mean squared error (MSE),  mean absolute deviation (MAD) or others. In the setting of a retailer, the observed quantity is the sales of individual products (denoted by $y$) and most machine learning approaches would then predict a single number $\hat{y}$ as a direct estimate of the sales. However, reducing the prediction to a single number does not allow to account for the uncertainty of the prediction or the dynamics of the system. Instead, it is imperative to predict the full PDF for each prediction to be able to optimize the subsequent operational decision. Unfortunately, most statistical or machine learning methods that predict full individual probability functions lack quantitative or at least qualitative evaluation methods to assess whether the full distribution has been forecast correctly, in particular in the tails of the distribution.

For an estimation of the determining parameters of an assumed functional form for the PDF, assessing the correctness of the PDF model output refers to the evaluation of the accuracy of the prediction of the different determining parameters. In the case of the negative binomial distribution used in this work, we have to verify that mean and variance are determined accurately, as well as checking that the choice of the underlying model can describe the observed data.

In the following, we will show how histograms of the CDF and inverse quantiles can be used to determine the accuracy of the predicted PDFs. Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a probability density distribution, even if the PDF is obtained empirically.

@@

As a first step, we use the inverse quantile function or probability integral transform (PIT), see e.g. \cite{Angus1994,casella2002statistical}. This offers a first qualitative assessment by looking at the distribution of the quantiles of the predicted density via the CDF. In case of a correctly calibrated predicted PDF, this distribution is expected to be uniform and any deviation can be interpreted as a hint that the predicted probability function is not correct \cite{diebold1998vevaluating}.

\subsubsection{Histogram of the CDF}
\label{sec:cdf_histo}

The CDF of a PDF $f(x)$ is defined as:

\begin{equation}
\label{eqn:CDF}
F_X(x) = P(X \le x) = \int_{-\infty}^{x} f_X(x^\prime) dx^\prime
\end{equation}

Here, $F_X(x)$ is the CDF with $\lim_{x \to -\infty}F_X(x) = 0$ and $\lim_{x \to \infty}F_x(x) = 1$. The cumulative distribution describes the probability that a the variable has a value smaller than $x$ and intuitively represents the area under $f(x^\prime)$ up to a point $x$.

If the CDF is continuous and strictly increasing, then the inverse of the CDF $F^{-1}(y)$ exists and is a unique real-valued number $x$ for each $y \in [0,1]$ so that we can write $F(x) = y$. The latter defines the inverse quantile function because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:

\begin{equation}
Q_\tau = F^{-1}(\tau)
\end{equation}

Using the example of the normal distribution with $\mathcal{N}(0,1)$ as shown in \fig \ref{fig:PdfCdf}, we can identify the median ($\tau = 0.5$) by first looking at the CDF in the lower part of the figure, look at $y=0.5$ on the $y$-axis and then identify the point on the $x$ axis for both the PDF $f(x)$ and the CDF $F(x)$ that correspond to the quantile $\tau$. In the case of the normal distribution, this is of course the central value at zero.

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/PdfCdf}
\caption{\label{fig:PdfCdf} Probability distribution function and cumulative distribution funtion of a normal distrubution.}
\end{center}
\end{figure}

We can then interpret the CDF as a new variable $s = F(t)$, meaning that $F$ becomes a transformation that maps $t$ to $s$, i.e. $F:t \to s$. Accordingly,  $\lim_{t \to -\infty}s(t) = 0$ and $\lim_{t \to \infty}s(t) = 1$ and $s$ can be intuitively interpreted as the fraction of the distribution of $t$ with values smaller than $t$ from the definition of the CDF. This implies that the probability distribution of $s$, $g(s)$, is constant in the interval $s\in [0,1]$ in which $s$ is defined and $s$ can be interpreted as the cumulative distribution of its own probability distribution:

\begin{equation}
s = G(s) = \int_{-\infty}^{s} g(s^\prime) ds^\prime
\end{equation}

In case of discrete probability functions such as the negative binomial function, the same argument still holds but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g. \cite[p. 54]{casella2002statistical}. In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the inverse quantiles or values of the CDF is filled using random numbers according to the intervals of the CDF. For example, if the sales of zero items accounts for 75\% of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram is randomly chosen in the interval $[0,0.75]$. Proceeding similarly for all other observed values, the resulting histogram of CDF values should again be uniform as in the case of a continuous PDF.

A histogram of the quantiles (or PIT values) of each prediction is therefore expected to be uniformly distributed in $[0,1]$ if the predicted probability distribution $f(x)$ is correctly calibrated.

This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of quantiles (or PIT values) for five different cases. If both the choice of the model and the model parameters are estimated correctly, we would expect the uniform distribution as shown in the last part of the figure. The first two plots of \fig \ref{fig:cdf_histos} indicate a bias in the variance estimation and the third and fourth plot indicate a bias in the mean
estimation.

@@

\textcolor{red}{Need to add some formulas.}

@@

\textcolor{red}{... predict an individual demand volume for different product-location-date combinations ...}

\textcolor{red}{... example showing full prediction and evaluation chain ...}

\textcolor{red}{... plot showing individual PDF example ...}

@@

\textcolor{red}{... plot showing cdf histo ...}

@@

\textcolor{red}{... plot showing invquant profile for mean prediction on X-axis ...}

@@

\textcolor{red}{... plot showing invquant profile for product groups on X-axis ...}

@@
